---
title: "Coastal Bird monitoring at Boston Harbor Islands NRA (BOHA)"
author: "*Compiled on `r Sys.time()` by `r Sys.info()['user']` using `r R.Version()$version.string`*"
output: 
  html_document:
    highlight: textmate
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    fig_caption: yes
---
```{r,intro, results='markup', echo=F, include=FALSE}
cat("\\newpage")
library(NETNCoastalBirds)
library(tidyverse)
library(knitr)
library(magrittr)

path<-"C:/Users/aweed/OneDrive - DOI/Documents/NETN Protocols/Coastal Birds/R/R package/NETNCoastalBirds/inst/extdata/"

suppressWarnings({importCBBData(path, zip_name= "CSV_Dataset_Export_20230119.zip", new_env = TRUE)
})

```
# **Background**
 
Type some description here.

\pagebreak

# **Double-crested Cormorant (*Phalacrocorax auritus*)**
## Incubation surveys

```{r, cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}
options(knitr.kable.NA = '')

incub <- SumIncubation(df= NULL,  time = "year")
dcco<- filter(incub,Species_Code == "DCCO")
PlotBirds(dcco, raw_count = T, stat = "max")
cat("\\newpage")
kable(dcco[,c("Island","time","value")] %>% spread(time, value), caption= "Annual counts of Double-crested Cormorant",padding =2)
```

```{r, cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}
cat("\\newpage")
PlotBirds(dcco, scale = "log", stat = "max")
```

## Nest Surveys
```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}
# create data summarizing all ground-based nest counts
nests <- SumNestSurveys(df= NULL,time= "year")
# subset by DCCO to plot
dcco2<-filter(nests, Species_Code == "DCCO")
PlotBirds(dcco2, raw_count= TRUE, legend = TRUE, var = c("Nests", "Chicks", "Eggs"))
```
\pagebreak

# **Great Black-backed Gull (*Larus marinus*)**
## Incubation surveys
```{r, cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12, message=FALSE}
options(knitr.kable.NA = '')
#incub <- SumIncubation(time = "year")
GBBG<- filter(incub,Species_Code == "GBBG")
PlotBirds(GBBG, raw_count = T, stat = "max")
cat("\\newpage")
kable(GBBG[,c("Island","time","value")] %>% spread(time, value), caption= "Annual counts of Great Black-backed Gull",padding =2)
```

```{r, cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}
cat("\\newpage")
PlotBirds(GBBG, scale = "log", stat = "max")
```

## Nest Surveys
```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12, message=FALSE}
gbbg2<-filter(nests, Species_Code == "GBBG")
PlotBirds(gbbg2, raw_count= TRUE, legend = TRUE, var = c("Nests", "Chicks", "Eggs"))
```
\pagebreak 

# **Herring Gull (*Larus argentatus*)**
## Incubation surveys
```{r, cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}
options(knitr.kable.NA = '')
#incub <- SumIncubation(time = "year")
HERG<- filter(incub,Species_Code == "HERG")
PlotBirds(HERG, raw_count = T, stat = "max")
cat("\\newpage")
kable(HERG[,c("Island","time","value")] %>% spread(time, value), caption= "Annual counts of Herring Gull",padding =2)
```

```{r, cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12, message=FALSE}
cat("\\newpage")
PlotBirds(HERG, scale = "log", stat = "max")
```

## Nest Surveys
```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12, message=FALSE}
herg2<-filter(nests, Species_Code == "HERG")
PlotBirds(herg2, raw_count= TRUE, legend = TRUE, var = c("Nests", "Chicks", "Eggs"))
```

# **Common Eider (*Somteria mollissima*)**
## Creche Surveys
### Current year surveys by date
```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}
options(knitr.kable.NA = '')
creche <- SumCreche(df = CBB_TABLES$qry_Dataset_4_Survey_Creche, time ="date")
PlotBirds(creche, year= "2022", raw_count = T, stat = "sum", legend= T, var= c("COEI Ducklings" ,"Adult female COEI tending"))
cat("\\newpage")
yr<-"2022"
df<- filter(creche, year== yr & !Island == "All Islands") %>% select(Island,time,variable,value) %>% spread(time,value) %>% mutate(variable= abbreviate(variable, minlength = 12, method= "both.sides"))
kable(df, caption = paste0("Creche surveys during ",yr)) 
```

### Annual counts over time by variable
```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12, message=FALSE}
creche_y <- SumCreche(df = CBB_TABLES$qry_Dataset_4_Survey_Creche,time ="year")
VAR<-unique(creche_y$variable); VAR<-set_names(VAR, nm=VAR)
plots<-purrr::map(VAR, ~PlotBirds(creche_y, var = . , raw_count = T , stat = "max", print= F, legend = F))
print(plots[[1]]);cat("\\newpage")
print(plots[[2]]);cat("\\newpage")
print(plots[[3]]);cat("\\newpage")
print(plots[[4]])
cat("\\newpage")
```

## Nest Surveys
```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}
coei<-filter(nests, Species_Code == "COEI")
PlotBirds(coei, raw_count= TRUE, legend = TRUE, var = c("Nests", "Chicks", "Eggs"))
```
\pagebreak

# **Black-crowned night Heron (*Nycticorax nycticorax*)**
## Nest Surveys
```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}
bcnh<-filter(nests, Species_Code == "BCNH")
PlotBirds(bcnh, raw_count= TRUE, legend = TRUE, var = c("Nests", "Chicks", "Eggs"))
```
\pagebreak 

# **Glossy Ibis (*Plegadis falcinellus*)**
## Nest Surveys
```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12, message=FALSE}
glib<-filter(nests, Species_Code == "GLIB")
PlotBirds(glib, raw_count= TRUE, legend = TRUE, var = c("Nests", "Chicks", "Eggs"))
```
\pagebreak 

# **Great Egret (*Ardea alba*)**
## Nest Surveys
```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}
greg<-filter(nests, Species_Code == "GREG")
PlotBirds(greg, raw_count= TRUE, legend = TRUE, var = c("Nests", "Chicks", "Eggs"))
```
\pagebreak 

# **Snowy Egret (*Egretta thula*)**
## Nest Surveys
```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}
sneg<-filter(nests, Species_Code == "SNEG")
PlotBirds(sneg, raw_count= TRUE, legend = TRUE, var = c("Nests", "Chicks", "Eggs"))
```
