---
title: "Coastal Bird monitoring at Boston Harbor Islands NRA (BOHA)"
author: "*Compiled on `r Sys.time()` by `r Sys.info()['user']` using `r R.Version()$version.string`*"
output: 
  pdf_document:
    toc: yes
    fig_caption: yes
  html_document:
    toc: true
    number_sections: false
---
```{r,intro, results='markup', echo=F, include=FALSE}
cat("\\newpage")
library(NETNCoastalBirds)
library(tidyverse)
library(knitr)
library(magrittr)
```
\pagebreak

# Background
 
Type some description here.

\pagebreak

# **Double-crested Cormorant (*Phalacrocorax auritus*)**
##Incubation surveys

```{r, cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 7.5,message=FALSE}
options(knitr.kable.NA = '')
incub <- SumIncubation(time = "year")
dcco<- filter(incub,Species_Code == "DCCO")
PlotBirds(dcco)
cat("\\newpage")
kable(dcco[,c("Island","time","value")] %>% spread(time, value), caption= "Annual counts of Double-crested Cormorant",padding =2)
```

```{r, cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 7.5,message=FALSE}
cat("\\newpage")
PlotBirds(dcco, scale = "log")
```

## Nest Surveys
```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 7.5,message=FALSE}
nests <- SumNestSurveys(time= "year", species= "DCCO")
PlotBirds(nests, caption = "no")
```
\pagebreak

# **Great Black-backed Gull (*Larus marinus*)**
##Incubation surveys
```{r, cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 7.5,message=FALSE}
options(knitr.kable.NA = '')
#incub <- SumIncubation(time = "year")
GBBG<- filter(incub,Species_Code == "GBBG")
PlotBirds(GBBG)
cat("\\newpage")
kable(GBBG[,c("Island","time","value")] %>% spread(time, value), caption= "Annual counts of Great Black-backed Gull",padding =2)
```

```{r, cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 7.5,message=FALSE}
cat("\\newpage")
PlotBirds(GBBG, scale = "log")
```

## Nest Surveys
```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 7.5,message=FALSE}
nests <- SumNestSurveys(time= "year", species= "GBBG")
PlotBirds(nests, caption = "no")
```
\pagebreak 

# **Herring Gull (*Larus argentatus*)**
##Incubation surveys
```{r, cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 7.5,message=FALSE}
options(knitr.kable.NA = '')
#incub <- SumIncubation(time = "year")
HERG<- filter(incub,Species_Code == "HERG")
PlotBirds(HERG)
cat("\\newpage")
kable(HERG[,c("Island","time","value")] %>% spread(time, value), caption= "Annual counts of Herring Gull",padding =2)
```

```{r, cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 7.5,message=FALSE}
cat("\\newpage")
PlotBirds(HERG, scale = "log")
```

## Nest Surveys
```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 7.5,message=FALSE}
nests <- SumNestSurveys(time= "year", species= "HERG")
PlotBirds(nests, caption = "no")
```

# Common Eider (*Smateria mollissima*)
## Creche Surveys
### Current Year surveys by date
```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 7.5,message=FALSE}
options(knitr.kable.NA = '')
creche <- CrecheSum(time ="date")
PlotBirds(creche, year= "2019")
cat("\\newpage")
yr<-"2019"
df<- filter(creche, year== yr & !Island == "All Islands") %>% select(Island,time,variable,value) %>% spread(time,value) %>% mutate(variable= abbreviate(variable, minlength = 12, method= "both.sides"))
kable(df, caption = paste0("Creche surveys during ",yr)) 
```

### Annual counts over time by variable
```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 7.5,message=FALSE}
creche_y <- CrecheSum(time ="year")
VAR<-unique(creche_y$variable); VAR<-set_names(VAR, nm=VAR)
plots<-map(VAR, ~PlotBirds(creche_y, var = . ,print= "no", caption= "no"))
print(plots[[1]]);cat("\\newpage")
print(plots[[2]]);cat("\\newpage")
print(plots[[3]]);cat("\\newpage")
print(plots[[4]])
cat("\\newpage")
```

## Nest Surveys
```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 7.5,message=FALSE}
nests <- SumNestSurveys(time= "year", species= "COEI")
PlotBirds(nests, caption = "no")
```
\pagebreak

#**Black-crowned night Heron (*Nycticorax nycticorax*)**
## Nest Surveys
```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 7.5,message=FALSE}
nests <- SumNestSurveys(time= "year", species= "BCNH")
PlotBirds(nests, caption = "no")
```
\pagebreak 

#**Glossy Ibis (*Plegadis falcinellus*)**
## Nest Surveys
```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 7.5,message=FALSE}
nests <- SumNestSurveys(time= "year", species= "GLIB")
PlotBirds(nests, caption = "no")
```
\pagebreak 

#**Great Egret (*Ardea alba*)**
## Nest Surveys
```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 7.5,message=FALSE}
nests <- SumNestSurveys(time= "year", species= "GREG")
PlotBirds(nests, caption = "no")
```
\pagebreak 

#**Snowy Egret (*Egretta thula*)**
## Nest Surveys
```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 7.5,message=FALSE}
nests <- SumNestSurveys(time= "year", species= "SNEG")
PlotBirds(nests, caption = "no")
```
