# Overview
title: "Coastal Breeding bird monitoring data package metadata workflow"
author: "Olivia Smith, Aaron Weed, Adam Kozlowski"
date: "9/16/2024"
output: xml

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
  
# Summary
This script acts as a template file for end-to-end creation of EML metadata in R for NPS DataStore. The XML metadata generated will be of sufficient quality for the Data Package Reference Type and can be used to automatically populate the DataStore fields for this reference type. The script utilizes multiple R packages and the inputs required are the current csv data files that comprise NETN's Water Quality & Quantity cumulative dataset.  

Generating a data package metadata file is a step by step process where each section (indicated by dashed lines) should be reviewed, edited if necessary, and run one at a time. After completing a section there is often something to do external to R (e.g. open a text file and add content). Several EMLassemblyline functions are decision points and may only apply to certain data packages. This workflow takes advantage of the NPSdataverse, an R-based ecosystem that includes external EML creation tools such as the R packages EMLassemblyline and EML. However, these tools were not designed to work with DataStore. Therefore, the NPSdatavers and this workflow also incorporate steps from NPS-developed R packages such as EMLeditor and DPchecker. You will necessarily over-write some of the information generated by EMLassemblyline. That is OK and is expected behavior.

# Good additional references
[EMLassemblyline](https://ediorg.github.io/EMLassemblyline/)
[EMLeditor](https://nationalparkservice.github.io/EMLeditor/index.html)
[NPS EML Script](https://nationalparkservice.github.io/NPS_EML_Script/)
[EVER Veg Map AA dataset](https://github.com/nationalparkservice/NPS_EML_Script/tree/main/Example_files) (for testing purposes)

# Install and Load R Packages 
Install packages. If you have not recently installed packages, please re-install them. You will want to make sure you have the latest versions of all the NPSdataverse packages - QCkit, EMLeditor, DPchecker, and NPSutils - as they are under constant development. If you are on the VPN, you will need to set your CRAN mirror to Texas 1. If you run into errors installing packages from github on NPS computers you may first need to run:
```{r renviron, eval=FALSE}
options(download.file.method="wininet")
```
```{r package_install, eval=FALSE}
#install packages
install.packages(c("devtools", "tidyverse", "NETNCoastalBirds"))
#load devtools
library(devtools)
library(tidyverse)
library(NETNCoastalBirds)
#install/update NPSdataverse
devtools::install_github("nationalparkservice/NPSdataverse")
```
##Troubleshooting updating NPSdataverse (optional)
Downloading the NPSdataverse from the NPS GitHub repo may be problematic due to API limits if you are not an authenticated user.  To increase your GitHub API rate limit you need to generate a Personal Access Token (PAT). Generating a PAT requires a GitHUb account. Quick version is below.  Or the entire process is nicely outlined in this script: https://gist.github.com/z3tt/3dab3535007acf108391649766409421
```{r GitHub_token, eval=FALSE}
# to create a personal access token for authentication:
#usethis::create_github_token()
# to set personal access token:
#credentials::set_github_pat("Enter_Your_PAT_Here")
```

```{r load_packages}
# When loading packages, you may be advised to update to more recent versions
# of dependent packages. Most of these updates likely are not critical.  
# However, it is important that you update to the latest versions of QCkit, 
# EMLeditor, DPchecker and NPSutils as these NPS packages are under constant 
# development.
library(NPSdataverse)
library(tidyverse)
```

## Set the overall package details
All of the following items should be reviewed and updated to fit the package you are working on. For lists of more than one item, keep the same order (i.e. item #1 should correspond to the same file in each list).

#### Dataset name
Custom variable added to make it easier to change the dataset file name once and have it cascade throughout this script. 



```{r dataset_filename}
# note: instead of hard coding the date we can append the date based on system date the code is run

dataset <- paste0("NETN_Coastalbirds_",Sys.Date())
```

#### Metadata filename
This becomes the file name of your .xml file. Be sure it ends in _metadata to comply with data package specifications. You do not need to include the extension (.xml).
```{r metadata_id}

metadata_id <- paste(dataset,"NETN_CoastalBirds_metadata", sep="")
```

#### Package Title
Give the data package a title. FAIR principles suggest titles of between 7 and 20 words. Be sure to make your title informative and consider how a naive user would interpret it.
```{r package_title}
package_title <- "Coastal Breeding bird monitoring data package for NETN and BOHA"
```

#### Data collection status
Choose from either "ongoing" or "complete"
```{r data_status}
data_type <- "ongoing"
```

#### Path to data file(s)
Tell R where your data files are. If they are in the working directory, you can set the working_folder to `getwd()`. If they are in a different directory you will need to specify that directory. Find the working directory with getwd() command.
```{r working_dir}
sites <- GetSites()
segments <- GetSegments()
events <- GetEvents()
incub <- GetIncubationData()
ground.nests <- GetNestData()
creche<-GetCrecheData()
AMOY<-GetAMOYData()
surv <- GetSurveillanceData()
incidental <- GetIncidentalData()
counts <- GetCountSum()
```

#### List data files
Tell R what your data files are called. 
```{r list_files}
data_files <- list(sites, segments, events, incub, ground.nests, creche, AMOY,
                   surv, incidental, counts)
```

#### Name the data files
These should be relatively short, but perhaps more informative than the actual file names. Make sure that they are in the same order as the files in data_files.
```{r data_names}
data_names <- c("Coastal bird survey sites",
                "Coastal bird survey route segments",
                "Coastal bird survey summary",
                "Coastal bird incubating adult data",
                "Coastal bird nest data",
                "Common Eider crèche data",
                "American Oystercatcher survey data",
                "Coastal bird surveillance survey data",
                "Coastal bird incidental observations",
                "Coastal bird end of year summary"
                )
```

#### Describe each data file. Data file descriptions should be unique and about 10 words long. Descriptions will be used in auto-generated tables within the ReadMe and DRR. If you need to use more than 10 words, consider putting that information into the abstract, methods, or additional information sections. Again, make sure these are in the same order as the files they are describing.
```{r data_descriptions}
data_descriptions <- c("Summary information for survey sites, including geographic data",
                       "Summary information for survey route segments for each site",
                       "Summary information for coastal bird surveys",
                       "Raw data for boat-based surveys of incubating coastal birds",
                       "Raw data for ground-based coastal bird nest surveys",
                       "Raw data for boat-based COEI crèche surveys",
                       "Raw data for AMOY surveys",
                       "Observations of target species detected outside of survey routes",
                       "Coastal bird observations of incidental (non-target) species",
                       "End of year summary data, includes non-NETN collected data"
                       ) 
```

#### Placeholder URL for data files
EMLassemblyline needs to know where the data files will be (a URL). However, because you have not yet initiated a draft reference in DataStore, it isn't possible to specify a URL. Instead, insert a place holder. Don't worry - this information will be updated later on when you add a Digital Object Identifier(DOI) to the metadata.
```{r temp_data_url}
data_urls <- c(rep("temporary URL", length(data_files)))
```

#### Taxonomic information
Specify where you taxonomic information is. This can be a single file or a list of files and fields (columns) with scientific names that will be used to automatically generate the taxonomic coverage metadata. We suggest using [DarwinCore](https://dwc.tdwg.org/terms) for column names, such as "scientificName". If your data package does not have taxonomic data, skip this step.
```{r taxonomic_info}
# the file(s) where scientific names are located:
data_taxa_tables <- data_files

# the column where your scientific names are within the data files.
data_taxa_fields <- "ScientificName"
```

#### Geographic information
Specify the tables and fields that contain geographic coordinates and site names. This information will be used to fill out the geographic coverage elements of the metadata. If your data package does not have geographic information skip this step. If the only geographic information you are supplying is the park units (and their bounding boxes) you can also skip this step; Park Units and the corresponding GPS coordinates for their bounding boxes will be added at a later step. If your coordinates are in UTMs and not GPS, try the [`convert_utm_to_ll()`](https://nationalparkservice.github.io/QCkit/reference/convert_utm_to_ll.html) function in the [QCkit package](https://nationalparkservice.github.io/QCkit/)
```{r geo_info}
data_coordinates_table <- data_files
data_latitude <- "Latitude"
data_longitude <- "Longitude"
data_sitename <- "Site_Name"
```

#### Temporal information
This should indicate collection date of the first and last data point in the data package (across all files) and does not include any planning, pre- or post-processing time. The format should be one that complies with the International Standards Organization's standard 8601. The recommended format for EML is: YYYY-MM-DD, where Y is the four digit year, M is the two digit month code (01 - 12 for example, January = 01), and D is the two digit day of the month (01 - 31). Using an alternate format or setting the date to the future will cause errors down the road!
```{r temp_info}
startdate <- ymd("2007-05-15")
enddate <- ymd("2024-07-08")
```

## EMLassemblyline Functions
The next set of functions are meant to be considered one-by-one and only run if applicable to a particular data package. The first year will typically see all of these run, but if the data format and protocol stay constant over time it may be possible to skip some in future years. Additionally some datasets may not have geographic or taxonomic component.

#### FUNCTION 1 - Core Metadata Information
This function creates blank .txt template files for the abstract, additional information, custom units, intellectual rights, keywords, methods, and personnel. Be sure to edit the personnel text file in Excel as it has columns. Remember that the role "creator" is required! EMLassemblyline will also warn you if you do not  include a "PI" role, but you can ignore the warning; this role is not required. Typically these files can be reused between years.

We encourage you to craft your abstract in a text editor, NOT Word. Your abstract will be forwarded to data.gov, DataCite, google dataset search, etc. so it is worth some time to carefully consider what is relevant and important information for an abstract. Abstracts must be greater than 20 words. Good abstracts tend to be 250 words or less. You may consider including the following information: The premise for the data collection (why was it done?), why is it important, a brief overview of relevant methods, and a brief explanation of what data are included such as the period of time, location(s), and type of data collected. Keep in mind that if you have lengthy descriptions of methods, provenance, data QA/QC, etc it may be better to expand upon these topics in a Data Release Report or similar document uploaded separately to DataStore.

Currently this function inserts a Creative Common 0 license. The CC0 license will need to be updated. However, to ensure that the licence meets NPS specifications and properly coincides with CUI designations, the best way to update the license information is during a later step using `EMLeditor::set_int_rights()`. There is no need to edit this .txt file.
```{r core}
#See instructions above.  This step is updated during a later step.
#template_core_metadata(path = csv_folder, 
#                       license = "CC0") # that '0' is a zero!
```

#### FUNCTION 2 - Data Table Attributes
This function creates an "attributes_datafilename.txt" file for each data file. This can be opened in Excel (we recommend against trying to update these in a text editor) and fill in/adjust the columns for attributeDefinition, class, unit, etc. refer to https://ediorg.github.io/EMLassemblyline/articles/edit_tmplts.html for helpful hints and `view_unit_dictionary()` for potential units. This will only need to be run again if the attributes (name, order or new/deleted fields) are modified from the previous year. NOTE that if these files already exist from a previous run, they are not overwritten.
```{r attributes}
#Before running, make sure only the data files (.csv) that you want to generate an attribute template for is in the working directory!!
template_table_attributes(path = working_folder, 
                          data.table = data_files, 
                          write.file = TRUE)
```

#### FUNCTION 3 - Data Table Categorical Variable
This function Creates a "catvars_datafilename.txt" file for each data file that has columns with a class = categorical. These .txt files will include each unique 'code' and allow input of the corresponding 'definition'.  NOTE that since the list of codes is harvested from the data itself, it's possible that additional codes may have been relevant/possible but they are not automatically included here. Consider your lookup lists carefully to see if additional options should be included (e.g if your dataset DPL values are all set to "Accepted" this  function will not include "Raw" or "Provisional" in the resulting file and you may want to add those manually). NOTE that if these files already exist from a previous run, they are not overwritten.
```{r cat_vars}
#Before running, make sure only the attribute template (.txt) that you want to generate a catvars template for is in the working directory!!
template_categorical_variables(path = working_folder, 
                               data.path = working_folder, 
                               write.file = TRUE)
```

#### FUNCTION 4 - Geographic Coverage
If the only geographic coverage information you plan on using are park boundaries, you can skip this step. You can add park unit connections using EMLeditor, which will automatically generate properly formatted GPS coordinates for the park bounding boxes.

If you would like to add additional GPS coordinates (such as for specific site 
locations, survey plots, or bounding boxes for locations within a park, etc) please do. 

This function creates a geographic_coverage.txt file that lists your sites as points as long as your coordinates are in lat/long. If your coordinates are in UTM it is probably easiest to convert them first or create the geographic_coverage.txt file another way (see [QCkit](https://nationalparkservice.github.io/QCkit/) for R functions that will convert UTM to lat/long).
```{r geo_cov}
template_geographic_coverage(path = working_folder, 
                             data.path = working_folder,
                             data.table = data_coordinates_table, 
                             lat.col = data_latitude, 
                             lon.col = data_longitude,
                             site.col = data_sitename, 
                             write.file = TRUE)
```

#### FUNCTION 5 - Taxonomic Coverage
This function creates a taxonomic_coverage.txt file if you have taxonomic data. Currently supported authorities are 3 = ITIS, 9 = WORMS, and 11 = GBIF. In the example below, the function will first try to find the scientific name at ITIS and if it fails will then look at GBIF. If you have lots of taxa, this could take some time to complete.
```{r tax_cov}
template_taxonomic_coverage(path = working_folder, 
                            data.path = working_folder, 
                            taxa.table = data_taxa_tables,
                            taxa.col = data_taxa_fields, 
                            taxa.authority = c(3,11),
                            taxa.name.type = 'scientific', 
                            write.file = TRUE)
```

## Create an EML File
Run this (it may take a little while) and see if it validates (you should see 'Validation passed'). It will generate an R object called "my_metadata". The function could alert you of some issues to review as. Run the function `issues()` at the end of the process to get feedback on items that might be missing or need attention. Fix these issues and then re-run the `make_eml()` function.
```{r make_eml}
my_metadata <- make_eml(path = working_folder,
               dataset.title = package_title,
               data.table = data_files,
               data.table.name = data_names,
               data.table.description = data_descriptions,
               data.table.url = data_urls,
               temporal.coverage = c(startdate, enddate),
               maintenance.description = data_type,
               package.id = metadata_id,
               return.obj = TRUE, 
               write.file = FALSE)
            #Note:  this step creates the EML file as a data frame; it is written to file during a later step.
            #If you wish to write the results to file at this step toggle write.file code line to TRUE.
```

## Check for EML validity 
This is a good point to pause and test whether your EML is valid.
if your EML is valid you should see the following (admittedly cryptic) result:
```
# [1] TRUE
# attr(,"errors")
# character(0)
```
if your EML is not schema valid, the function will notify you of specific problems you need to address. We HIGHLY recommend that you use the EMLassemblyline and/or EMLeditor functions to fix your EML and do not attempt to edit it by hand.

```{r validate}
eml_validate(my_metadata)
```

## Add NPS specific fields to EML 
Now that you have valid EML metadata, you need to add NPS-specific elements and fields. For instance, unit connections, DOIs, referencing a DRR, etc. More information about these functions can be found at: [https://nationalparkservice.github.io/EMLeditor/](https://nationalparkservice.github.io/EMLeditor/).

#### Add Controlled Unclassified Information (CUI) codes
This is a required step. It is important to indicate not only that your data 
package contains CUI, but also to inform users if your data package does NOT
contain CUI because empty fields can be ambiguous (does it not contain CUI or
did the creators just miss that step?). You can choose from one of five CUI
dissemination codes. Watch out for the spaces! These are:
  * PUBLIC - Does NOT contain CUI.
  * FED ONLY - Contains CUI. Only federal employees should have access (similar to "internal only" in DataStore).
  * FEDCON - Contains CUI. Only federal employees and federal contractors should have access (also very much like current "internal only" setting in DataStore).
  * DL ONLY - Contains CUI. Should only be available to a named list of individuals (where and how to list those individuals TBD)
  * NOCON - Contains CUI. Federal, state, local, or tribal employees may have access, but contractors cannot.
More information about these codes can be found at: [https://www.archives.gov/cui/registry/limited-dissemination](https://www.archives.gov/cui/registry/limited-dissemination)

```{r cui_dissemination}
my_metadata <- set_cui_code(my_metadata, "PUBLIC")
# note that in this case the CUI code has been added to the original R object, 
# "my_metadata" but by giving it a new name, i.e. "my_meta2" you can
# create a new R object. Sometimes creating a new R object is preferable 
# because if you make a mistake you don't need to start over again.
```

## Intellectual Rights
EMLassemblyine and ezEML provide some attractive looking boilerplate for setting the intellectual rights. It looks reasonable and so is easy to just keep. However, NPS has some specific regulations about what can and cannot be in the intellectualRights tag. Use `set_int_rights()` to replace the text with NPS-approved text. Note: You must first add the CUI dissemination code using `set_cui()` as the dissemination code and license must agree. That is, you cannot give a data package with a PUBLIC dissemination code a "restricted" license (and vise versa: a restricted data package that contains CUI cannot have a public domain or CC0 license). You can choose from one of three options:
  * "restricted": If the data contains Controlled Unclassified Information (CUI), the intellectual rights must read: **"This product has been determined to contain Controlled Unclassified Information (CUI) by the National Park Service, and is intended for internal use only. It is not published under an open license. Unauthorized access, use, and distribution are prohibited."**

  * "public": If the data do not contain CUI, the default is the public domain. The intellectual rights must read: **"This work is in the public domain. There is no copyright or license."**

  * "CC0": If you need a license, for instance if you are working with a partner organization that requires a license, use CC0: **"The person who associated a work with this deed has dedicated the work to the public domain by waiving all of his or her rights to the work worldwide under copyright law, including all related and neighboring rights, to the extent allowed by law. You can copy, modify, distribute and perform the work, even for commercial purposes, all without asking permission."**

The `set_int_rights()` function will also put the name of your license in a field in EML for DataStore harvesting.

```{r int_rights}
# choose from "restricted", "public" or "CC0" (zero), see above:
my_metadata <- set_int_rights(my_metadata, "public")
```

#### Add a data package DOI 
Add your data package's Digital Object Identifier (DOI) to the metadata. The `set_datastore_doi()` function requires that you are logged on to the VPN. It initiates a draft data package reference on DataStore, and populates the reference with a title pulled from your metadata, “[DRAFT] : <your data package title>”. This temporary title is purely for your tracking purposes and can easily be updated later. The `set_datastore_doi()` function will then insert the corresponding DOI for your data package into your metadata. Now that a draft reference has been initiated on DataStore, it is possible to fill in the online URL for each data file. `set_datastore_doi()` automatically does that for you too.  Here is the EMLeditor Function Reference documentation:  https://nationalparkservice.github.io/EMLeditor/reference/index.html

There are a few things to keep in mind:
  1) Your DOI and the data package reference are not yet active and are not publicly accessible until after review and activation/publication.
  2) We suggest you NOT manually upload files. If you manually upload your files, be sure to upload your data package to the correct draft reference! It is easy to create several draft references with the same draft title but different DataStore reference IDs and DOIs. So check the reference ID number carefully
  3) There is no need to fill in additional fields in DataStore at this point - many of them will be auto-populated based on the metadata you upload. Any fields you do populate will be over-written by the content in your metadata.
  
```{r DS_doi}
#set_datastore_doi():  Initiates a draft reference and inserts the reserved DOI into the metadata draft
#my_metadata <- set_datastore_doi(my_metadata)

#set_doi():  Check and set a DOI in the metadata draft (does not create a draft DataStore reference)
#example: my_metadata <- set_doi(my_metadata, "insert existing DOI")
my_metadata <- set_doi(my_metadata, "2305911")

#set_data_urls(): Update (or add) data table URLs
# For data packages on DataStore, no url is necessary:
#my_metadata <- set_data_urls(my_metadata)

# If data files are NOT on (or going to be on) DataStore, you must supply their location:
#my_metadata <- 3set_data_urls(my_metadata, "https://my_custom_repository.com/data_files")

#Full EMLeditor Function Reference documentation here:  https://nationalparkservice.github.io/EMLeditor/reference/index.html
```

#### Add information about a DRR (optional) 
If you are producing (or plan to produce) a DRR, add links to the DRR describing the data package.

You will need the DOI for the DRR you are drafting as well as the DRR's Title. Go to DataStore and initiate a draft DRR, including a title. For the purposes of the data package, there is no need to populate any other fields. At this point, you do not need to activate the DRR reference and, while a DOI has been reserved for your DRR,
it will not be activated until after publication so that you have plenty of 
time to construct the DRR.
```{r DRR_links}
#No DRR for Landbirds; skip this step
#my_metadata <- set_drr(my_metadata, 7654321, "DRR Title")
```

#### Set the language 
This is the human language (as opposed to computer language) that the data package and metadata are constructed in. Examples include English, Spanish, Navajo, etc. A full list of available languages is available from the Libraryof Congress. Please use the "English Name of Language" as an input. The function will then convert your input to the appropriate 3-character ISO 639-2 code.Available languages: [https://www.loc.gov/standards/iso639-2/php/code_list.php](https://www.loc.gov/standards/iso639-2/php/code_list.php)
```{r language}
my_metadata <- set_language(my_metadata, "English")
```
#### Add content unit links 
These are the park units where data were collected from, for instance ROMO, not ROMN. If the data package includes data from more than one park, they can all be listed. For instance, if data were collected from all park units within a network, each unit should be listed separately rather than the network. This is because the geographic coordinates corresponding to bounding boxes for each park unit listed will automatically be generated and inserted into the metadata. Individual park units will be more informative than the bounding box for the entire network.
```{r content_units}
park_units <- c("ACAD" , "APCO" , "BOWA" , "COLO" , "ELRO" , "FRSP" , "GETT" , "HOFR" , "HOFU" , "MABI" , "MIMA" , "MORR" , "PETE" , "RICH" , "SAGA" , "SAIR" , "SARA" , "VAFO" , "VAMA" , "WEFA")
my_metadata <- set_content_units(my_metadata, park_units)
```

#### Add the Producing Unit(s)
This is the unit(s) responsible for generating the data package. It may be a single park (ROMO) or a network (ROMN). It may be identical to the units listed in the previous step, overlapping, or entirely different.
```{r prod_units}
# a single producing unit:
#my_metadata <- set_producing_units(my_metadata, "MIDN")

# alternatively, a list of producing units:
my_metadata <- set_producing_units(my_metadata, c("MIDN", "NCBN", "NETN"))
```

## Validate your EML 
Almost done! This is another great time to validate your EML and make sure
Everything is schema valid. Run:
```{r val_eml}
eml_validate(my_metadata)
```
if your EML is valid you should see the following (admittedly cryptic):
# [1] TRUE
# attr(,"errors")
# character(0)

if your EML is not schema valid, the function will notify you of specific problems you need to address. We HIGHLY recommend that you use the EMLassemblyline and/or EMLeditor functions to fix your EML and do not attempt to edit it by hand.

## Write your EML to an xml file 
Now it's time to convert your R object to an .xml file and save it. Keep in mind that the file name should end with "_metadata.xml".
```{r write_eml}
#write_eml(my_metadata, "20240814_MIDN_NCBN_NETN_Landbird_metadata.xml")
write_eml(my_metadata, paste(dataset, "_metadata.xml", sep=""))
```

## Check your data package 
If your data package is now complete, you can run some test prior to upload to make sure that the package fits some minimal set of requirements and that the data and metadata are properly specified and coincide. This assumes that your data package is in the root of your R project. 
```{r congruence_tests}
# this assumes that the data package is the working directory
run_congruence_checks(skip_cols = "ExportDate")
#ExportDate column will be outside Start and End dates; so skip that test

# if your data package is somewhere else, specify that:
# run_congruence_checks("C:/Users/<yourusername>/Documents/my_data_package")
```

## Upload your data package 
If everything checked out, you should be ready to upload your data package! We recommend using `upload_data_package()` to accomplish this. The function automatically checks your DOI and uploads to the correct reference on DataStore. All of your files for the data package need to be in the same folder, there can be only one .xml file (ending in "_metadata.xml") and all the other files should be data files in .csv format. Each individual file should be < 32Mb.  If you have files > 32Mb, you will need to upload them manually using the web interface on DataStore. 
```{r upload}
# this assumes your data package is in the current working directory
upload_data_package()

# If your data package is somewhere else, specify that:
#upload_data_package("C:/Users/<yourusername>/Documents/my_data_package)
```

From within DataStore you should be able to extract all the information from the metadata file to populate the relevant DataStore fields.

Please don't activate your reference just yet! Data package references need to be reviewed prior to activation. We are still working on what that review process will look like.
