---
output:
  html_document:
    
    anchor_sections: FALSE
    
    
params:
  year: 2025
---


```{r,intro, results='hide', echo=F, include=FALSE}

library(NETNCoastalBirds)
library(tidyverse)
library(knitr)
library(magrittr)
library(kableExtra)
library(broom)
library(fuzzyjoin)

load("../data/tlu_Species.rda")
load("../data/SurveyEffortBySpecies.rda")
source("../docs/QAQCReport_functions.R")



#path<-"C:/Users/aweed/OneDrive - DOI/Documents/NETN Protocols/Coastal Birds/R/R package/NETNCoastalBirds/inst/extdata/"

#suppressWarnings({importCBBData(path, zip_name= "CSV_Dataset_Export_20230119.zip", new_env = TRUE)})

# plotting function to examine current vs historical counts


```

## Coastal Bird monitoring at Boston Harbor Islands (BOHA) {.tabset .tabset-fade .tabset-pills}

`r paste0("Compiled on ", Sys.time(), " by NETN using ",  R.Version()$version.string)`

### **Annual Gull and Cormorant Incubation Surveys** {.tabset}

#### Review Survey Effort in `r params$year`

**Staff should evaluate the tables to ensure that at least one incubation survey was completed per year and that at least one survey was designated as the Primary Survey (expert observation) for each Island and Route Segment combination. Note that there may be >1 Primary Survey for each Island/Event combination.**

``` {r incub_events, cache=T,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}

### Subset by outer Islands 
  
  outer<- c("Calf", "Little Calf", "Green", "The Graves", "Middle Brewster", 
            "Outer Brewster", "Shag Rocks","Little Brewster") 

incub_raw = GetIncubationData()

# surveys per segment 
incub_surveys<- incub_raw %>% filter(Species_Code == "DCCO" & year %in% params$year & Island %in% outer) %>% select( Island, Segment, Date, Nests = Unit_Count, Observer, Obs_Notes, Survey_Complete, Survey_Primary)

# Count the number of unique observers for primary and secondary incubation surveys by event
obs_incub<-incub_surveys %>% group_by(Island, Segment, Date, Survey_Complete, Survey_Primary) %>% tally(name= "Observers") %>% 
  # mutate(Observers = cell_spec(Observers, format= "html", background = ifelse(Survey_Primary =="Yes" & Observers < 1 , "orange", ""),
  #                           bold = ifelse(Survey_Primary =="Yes" & Observers >= 1, TRUE, FALSE))) %>% 
  spread(Date, Observers, fill = "No Survey")


# output Table 1 for Primary  surveys with conditional formatting 

kableExtra::kbl(filter(obs_incub,Survey_Primary == "Yes"), format = 'html', escape = FALSE, caption= paste0("Table 1.  Number of unique observers conducting a Primary Incubation Survey in ", params$year, " per date among Outer Islands. Confirm that at least one primary survey was conducted at all islands per date. Note that there may be an inconsistent nnumber of surveys conducted at Shag Rocks and the Graves due to the tidal stage at time of the survey.")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)


# output Table 2 for Non-Primary surveys with conditional formatting 

kableExtra::kbl(filter(obs_incub,Survey_Primary == "No"), 
          
                format = 'html', escape = FALSE, caption= paste0("Table 2.  Number of unique observers conducting a Non_Primary Incubation Survey in ", params$year, " per date among Outer Islands.")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)


# Count the number of unique observers by incubation survey event

events_incub<-incub_surveys %>% group_by(Island, Segment,Survey_Complete, Date, ) %>% tally(name= "Observers") %>% spread(Date, Observers, fill = "No Survey")
                  
# # Output Table 3 
# kableExtra::kbl(events_incub, escape = FALSE, align = 'c', caption= paste0("Table 3. Total number of unique observers (primary and non-primary) by survey date during Incubation Surveys conducted among Outer Islands in ", params$year, ".")) %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)

```

\pagebreak

#### DCCO {.tabset}

##### Raw counts from `r params$year`

`r Incubation_methods`

<!--

NETN staff should check the following sections for DCCO, GBBG, and HERG to ensure that:

**1. There is at least one completed primary survey (Survey_Complete = Yes) recorded for all island-segments in the database for each species and that the survey route (and count) was actually completed when you have records also denoting an incomplete survey was conducted on the same date.**

**2. There is consistency in the observer's name denoted during the primary surveys for each Island-Segment and the entire Outer Island boat route for each day and among surveys counted on other days in the same year. For instance, in some years (e.g., 2024) there may be >1 completed survey for each island-segment on different dates. NETN staff should check to confirm with the field team whether the repeated surveys were conducted by the same observer(s) among dates if denoted otherwise in the database. If so, the initials of the Principal Investigator should be denoted for all Primary Survey records in the database.**

**3. All field observational notes (Obs_Notes) are clearly understood and the ramifications on the survey counts are reconciled in the database.**

**4. Incubating adult nest counts during the current year are reasonable compared to the long term record .**

  **a. In a year when repeated primary surveys are conducted either by separate teams or on different dates, NETN staff should discuss     variation in nest counts among events to determine whether one event should be considered primary over the other and whether calculating     the average across the events represents the best possible estimate for nesting in that year.**

  **b. NETN staff should examine graphical and tabular summaries of the long term survey record to assess whether the current year count    is an outlier. See section: Nesting per island over time for each species.**

--->

```{r, cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 12, fig.width= 12,message=FALSE}

PlotWeeklyObserver(survey = "incubation", species = "DCCO", current_yr = params$year)


dcco_incub<- incub_raw %>% filter(Species_Code == "DCCO" & year %in% params$year & Survey_Primary == "Yes") %>% select( Island, Segment, Date, Survey_Complete, Nests = Unit_Count, Observer,  Obs_Notes)

# raw count data by observer
kableExtra::kbl(dcco_incub, format = "html",  caption= "Table 3. Current year counts of Double-crested Cormorant from primary surveys conducted among Outer Islands",  padding =2) %>%
  kable_styling(bootstrap_options = c("condensed"), fixed_thead = T)

```

##### Nesting over time

`r SumIncubation_methods`


```{r, cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 12, fig.width= 12,message=FALSE}

options(knitr.kable.NA = '')

incub <- SumIncubation(  time = "year") 
# %>% mutate(shape = case_when(
#   year == params$year ~ "8",
#   .default = "19" )) %>% 
#   dplyr::mutate(color = case_when(
#   year == params$year ~ "red",
#   .default = "black" ))


DCCO<- filter(incub,Species_Code == "DCCO")

# add table of raw counts
kableExtra::kbl(DCCO[,c("Island","time","value")] %>% spread(time, value), caption= "Table 4. Maximum annual counts of Double-crested Cormorant",padding =2) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)

cat("\\newpage")

# plot time series per Island
PlotBirds(DCCO, raw_count = T, stat = "max", plot_title = FALSE)

# create violin plot

plt_var(type= "violin", survey = "incubation", species = "DCCO", var ="Incubating Adults")


```


#### GBBG {.tabset}

##### Raw counts from `r params$year`

`r Incubation_methods`

```{r, cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 12, fig.width= 12,message=FALSE}

PlotWeeklyObserver(survey = "incubation", species = "GBBG", current_yr = params$year)

gbbg_incub<- incub_raw %>% filter(Species_Code == "GBBG" & year %in% params$year & Survey_Primary == "Yes") %>% select( Island, Segment, Date, Survey_Complete,Nests = Unit_Count, Observer,  Obs_Notes)

# raw count data by observer
kable(gbbg_incub, caption= "Table 5. Current year counts of Great black-backed Gulls from primary surveys",  padding =2) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)

```

##### Nesting over time

`r SumIncubation_methods`


```{r, cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 12, fig.width= 12, message=FALSE}
options(knitr.kable.NA = '')
#incub <- SumIncubation(time = "year")
GBBG<- filter(incub,Species_Code == "GBBG")

# add table of raw counts
kable(GBBG[,c("Island","time","value")] %>% spread(time, value), caption= "Table 6. Maximum annual counts of Great Black-backed Gull",padding =2) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)

# plot time series per Island
PlotBirds(GBBG, raw_count = T, stat = "max")
cat("\\newpage")

# create violin plot

plt_var(type= "violin", survey = "incubation", species = "GBBG", var ="Incubating Adults")

```
\pagebreak 

#### HERG {.tabset}

##### Raw counts from `r params$year`

`r Incubation_methods`

```{r, cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 12, fig.width= 12,message=FALSE}

PlotWeeklyObserver(survey = "incubation", species = "HERG", current_yr = params$year)

herg_incub<- incub_raw %>% filter(Species_Code == "HERG" & year %in% params$year & Survey_Primary == "Yes") %>% select( Island, Segment, Date, Survey_Complete, Nests = Unit_Count, Observer, Survey_Complete, Obs_Notes)

# raw count data by observer
kable(herg_incub, caption= "Table 7. Current year counts of Herring Gulls from primary surveys",  padding =2) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)

```

##### Nesting over time

`r SumIncubation_methods`

```{r, cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 12, fig.width= 12,message=FALSE}
options(knitr.kable.NA = '')
#incub <- SumIncubation(time = "year")

HERG<- filter(incub,Species_Code == "HERG")

# add table of raw counts
kable(HERG[,c("Island","time","value")] %>% spread(time, value), caption= "Table 8. Maximum annual counts of Herring Gull",padding =2) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)


# plot time series per Island
PlotBirds(HERG, raw_count = T, stat = "max")
cat("\\newpage")

# create violin plot

y<- plt_var(type= "violin", survey = "incubation", species = "HERG", var ="Incubating Adults")

```

### **Annual COEI Creche Surveys** {.tabset}



#### Review Survey Effort in `r params$year`

`r Crechmethods`

**NETN staff should evaluate the below tables to ensure that (1) at least 3 to 5 creche surveys were completed per year during the peak season by an expert birder (Primary Survey), (2) evaluate, document and reconcile why there may be incomplete surveys denoted as a primary survey for a particular date, and (3) confirm that counts for all four life stages were collected at each Island-Segment and Date.**


``` {r creche_events, cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}

creche_raw<-GetCrecheData()

# filter out complete surveys per segment for current year 
creche_surveys<- creche_raw %>% filter(year %in% params$year) %>% select(Island, Segment, Date, Species_Unit, Observer, Group_Notes, Survey_Notes, Survey_Complete, Survey_Primary)

# create tables summarizing creche survey events by observer
events_creche<-creche_surveys %>% select(-Group_Notes, -Survey_Notes, -Species_Unit) %>%  group_by(Island, Segment, Date, Survey_Complete,  Survey_Primary) %>% distinct() %>%  tally(name= "Observers") %>% spread(Date, Observers, fill ="No survey")

# Table of Primary Surveys
kable(events_creche %>% filter(Survey_Primary == "Yes") %>% select(-Survey_Primary), caption= paste0("Table 9. Number of unique observers conducting complete, Primary Surveys by date in ", params$year ," . At least one observer should be denoted for each Island-Segment date unless the area is not visible during high tide (Roaring Bulls, The Graves, Shagh Rocks).")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)

# Table of Non-Primary Surveys
kable(events_creche %>% filter(Survey_Primary == "No") %>% select(-Survey_Primary), caption= paste0("Table 10. Number of unique observers conducting complete, Non-Primary Surveys by date in ", params$year)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)


# create table summarizing number of life stages recorded per event (island and date) for only complete surveys
obs_creche<-creche_surveys  %>% filter(Survey_Primary == "Yes") %>% select(-Group_Notes, -Observer, -Survey_Notes) %>% group_by(Island, Segment, Date, Species_Unit) %>%  distinct() %>% group_by(Island, Segment, Date) %>% tally(name= "Life_Stages") %>% 
  spread(Date, Life_Stages, fill ="No survey")

#obs_creche$Life_Stages<-cell_spec(obs_creche$Life_Stages, color= ifelse(obs_creche$Life_Stages == 4, "black", "red"))

kable(obs_creche, caption= paste0("Table 11. Number of life stages counted during complete, Primary Surveys by date in ", params$year ,". Four life stages should be monitored per event.")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)
cat("\\newpage")
```

The following records were denoted as **incomplete surveys** in `r params$year`:
 
``` {r creche_rev, cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}

if(nrow( creche_surveys %>% filter(Survey_Complete == "No" & Survey_Primary == "Yes")) <1 ){
  print(paste0("There are no records denoted as Incomplete in ", params$year))
  
} else{

kable(creche_surveys %>% filter(Survey_Complete == "No" & Survey_Primary == "Yes"), caption = "Table 12. Details from records listed as the Primary Survey but also as Incomplete") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)
}
```


#### Review surveys by date for (`r params$year`) {.tabset}

`r SumCreche_Daily`

```{r crecheday,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}
options(knitr.kable.NA = '')

creche <- SumCreche(time ="date", islands = "outer") # Summarize daily Primary Surveys

# plot counts of F-tending and ducklings by survey date in current year for each island
y1<- PlotBirds(filter(creche,!variable == "Average creche size"), year= params$year, raw_count = T, stat = "sum", legend= T, print= FALSE)

(y1 + scale_color_manual(values = c("black","darkgray", "blue"))+ labs(title= "Daily counts of Common Eider (COEI) per lifestage and  Island"))

```

##### Ducklings

`r CrechmethodsHist`

```{r ducks,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 12, fig.width= 12,message=FALSE}
options(knitr.kable.NA = '')

creche <- SumCreche(time ="date", islands = "outer") # Summarize daily Primary Surveys

# plot current year counts per life stage compared to historical data for all island segments

VAR<-c("Chick", "F-Lone", "F-Tend"); VAR<-set_names(VAR, nm=VAR)

# per island plots by life stage (variable)
COEIplots<-purrr::map(VAR, ~PlotWeeklyObserver(survey = "creche", current_yr = params$year, var = ., print= FALSE ))

COEIplots[[1]]

# Output table of raw counts per lifestage

df<- filter(creche, year== params$year & !Island == "All Islands" ) %>% select(Island,Date = time,variable,value) %>% spread(Date,value) 

kable(df %>% filter(variable == "COEI Ducklings") , caption = paste0("Counts per life stage of Common Eider during Primary Surveys ", params$year))  %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)

cat("\\newpage")

```

##### Females tending ducklings

`r CrechmethodsHist`

```{r Ftend,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 12, fig.width= 12,message=FALSE}
options(knitr.kable.NA = '')

COEIplots[[3]]


kable(df %>% filter(variable == "Adult female COEI tending") , caption = paste0("Counts per life stage of Common Eider during Primary Surveys ", params$year))  %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)

cat("\\newpage")

```

##### All  Females

`r CrechmethodsHist`

```{r Ftotal,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 12, fig.width= 12,message=FALSE}
options(knitr.kable.NA = '')

COEIplots[[2]]

kable(df %>% filter(variable == "Total Number of Female COEI Observed") , caption = paste0("Counts per life stage of Common Eider during Primary Surveys ", params$year))  %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)

```


#### Annual counts over time by life stage
`r SumCreche_Annual`

```{r creche_year,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12, message=FALSE}
creche_y <- SumCreche(df = NULL ,time ="year") %>% 
  mutate(shape = case_when(
  year == params$year ~ "8",
  .default = "19" )) %>% 
  mutate(color = case_when(
  year == params$year ~ "red",
  .default = "black" ))

VAR<-unique(creche_y$variable); VAR<-set_names(VAR, nm=VAR)

# per island plots by life stage (variable)
plots<-purrr::map(VAR, ~PlotBirds(creche_y, var = . , raw_count = T , stat = "max", print= F, legend = F))

# All island plots by life stage (variable)
plots2<-purrr::map(VAR, ~PlotBirds(creche_y, var = . , island ="All Islands", raw_count = T , stat = "max", print= F, legend = F, facet= "variable", plot_title = F))

# plot F-tend stats
print(plots[[1]]);cat("\\newpage")# F-tend time series per island
print(plots2[[1]]);cat("\\newpage")# F-tend time series all islands

# plot duckling stats

print(plots[[3]]);cat("\\newpage")# Ducklings time series per island
print(plots2[[3]]);cat("\\newpage")# Ducklings time series all islands

# plot all female counts
print(plots[[4]]);cat("\\newpage")# total femalestime series per island
print(plots2[[4]]);cat("\\newpage")# total females for all islands


# create box plot of all life stages for all islands
ggplot(data= filter(creche_y, Island == "All Islands", stat == "max", !variable == "Average creche size"), aes(x= variable, y= value, fill= variable, group= variable)) +
  geom_violin(alpha= 0.3, notch= F)+ 
  geom_jitter(aes(shape = shape, color= color), size= 3) +
  scale_color_manual(values = c("black", "red"))+ 
  coord_flip()+ facet_wrap(~variable, scales = "free_x")+
  scale_shape_manual(values = c(19, 8))+ 
  theme_bw()+
  theme(legend.position = "none")+
  labs(y = "Count", x ="")+
  ggtitle(paste0("Variation in creche counts per life stage from 2007 to ", params$year, " for Outer Islands. The red asterisks denotes current year count."))+
  theme(axis.text.y = element_text(color="black", vjust= 0.5, size = 16)) +
         theme(axis.text.y = element_text(angle = 0,  vjust=0,size = 12 )) +
        theme(axis.text.y =element_blank())  +
         theme(strip.text.x= element_text(size=12, face=c("bold.italic"))) +
         theme(axis.title.x =element_text(size = 16, face ="bold", vjust= 0, debug=F)) +
         theme(axis.title.y =element_text(size = 16, face ="bold", vjust= 1, debug=F)) +
         theme(panel.background =  element_rect(fill="white", colour="black")) +
         theme(panel.grid.major = element_line(colour = "grey90")) +
         theme(plot.title=element_text(size=12, vjust=2, face= "bold")) +
         theme(strip.background= element_rect(size=10, color="gray" ))

# assess trends in ducklings and f-tend

df<- filter(creche_y, Island == "All Islands" & !variable == "Average creche size" & stat =="max" & year %in% c(2007:2025))

PlotBirds(df, island ="All Islands", raw_count = T , stat = "max", print= F, legend = F, facet= "variable", plot_title = F) + geom_smooth(method = "lm")+ scale_color_manual(values = c("black","darkgray", "blue"))


  # fit linear trend models 

ducklings<-lm(value ~ year, offset = Survey_Size, data = filter(creche_y, variable =="COEI Ducklings")) %>% tidy()

kable(ducklings, caption = "Linear Trend Model Summary for Ducklings", digits = 3)


ftend <-lm(value ~ year, offset = Survey_Size, data = filter(creche_y, variable == "Adult female COEI tending")) %>% tidy()

kable(ftend, caption = "Linear Trend Model Summary for Tending Females", digits = 3)

FTotal <-lm(value ~ year, offset = Survey_Size, data = filter(creche_y, variable == "Total Number of Female COEI Observed")) %>% tidy()

kable(FTotal, caption = "Linear Trend Model Summary for All Females", digits = 3)


```
\pagebreak

#<!--

### **Ground-based Nest Surveys** {.tabset}

The following tables and graphs summarize data from boat-based creche surveys in the program's database conducted in `r params$year`. 

**NETN staff should evaluate the below tables for accuracy and completeness to ensure that (1) at least one ground survey was completed during the peak season for all appropriate Island-Segments (see protocol), (2) evaluate, document and reconcile why there may be incomplete surveys denoted as a primary survey for a particular date, (3) confirm that at least one survey was designated as the Primary Survey for each species and per date, and (4) evaluate what proportion of surveys conducted direct vs indirect (flush) counting methods.**

#### **Review Survey Effort in `r params$year`**

``` {r ground_events, cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}

# Notes: 
# 
# 1. It may be useful to bring in the full table of islands that were supposed to be surveyed to check completeness of surveying work
# 2. Double check that partial surveys occurred and that the portion of the route survey was estimated in the DB. As of now the % surveyed is not being imported into the R package object (confirm data field with Adam).



# bring in raw event level data

nests_raw<-GetNestData()

# create data summarizing all ground-based nest counts
nests <- SumNestSurveys(df= NULL,time= "year") %>% mutate(shape = case_when(
  year == params$year ~ "8",
  .default = "19" )) %>% 
  mutate(color = case_when(
  year == params$year ~ "red",
  .default = "black" ))

# surveys per segment 
ground_surveys<- nests_raw %>% filter(year %in% params$year) %>% select(Species_Code, Island, Segment, Date, Nests = Unit_Count, Observer, Obs_Notes, Survey_Complete, Survey_Primary)

  if(nrow(ground_surveys) == 0){
    paste0("No ground-based nest surveys were collected during ", params$year)
    
    }else {

# create table showing survey route completion status by species and date
obs_ground<-ground_surveys %>% select(-Observer, -Obs_Notes, -Nests) %>% distinct() %>% spread(Species_Code, Survey_Complete) %>% arrange(Date, Island, Segment)

kable(obs_ground, caption= paste0("Table 13. Status of ground-based surveys by species and island in ", params$year ,". Cell values denotes whether or not the survey was fully completed. No entry denotes that no survey was conducted.")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)

# create table showing how nests were counted vs estimated for each species, to be subset below under each species
# this code filters out nest data for the current year, if there were no ground based counts in the current survey year this will be denoted

ground_methods<- nests_raw %>% filter(year %in% params$year) %>% select(Species_Code, Island, Segment, Date, Nests = Unit_Count, Nest_Status, Survey_Complete, Survey_Primary) %>% group_by(Species_Code, Island, Segment, Date, Survey_Complete, Survey_Primary, Nest_Status) %>% summarise(Nests = sum(Nests)) %>% spread(Nest_Status, Nests, fill =0) %>% rowwise() %>% group_by(Species_Code, Island, Segment, Date) %>% mutate(Total = sum(Abandoned,	Depredated,	Estimate,	Fledged,	Normal,	Other,	Unknown	)) %>% arrange(Date, Island, Segment)
  
  
  ground_notes<- nests_raw %>% filter(year %in% params$year ) %>% group_by(Species_Code, Island, Segment, Date) %>% 
  mutate(c_Survey_Notes = paste0(unique(Survey_Notes), collapse = ";")) %>% distinct() %>% select(Species_Code, Island, Segment, Survey_Complete, Date, c_Survey_Notes)  %>% arrange(Date, Island, Segment)

ground_notes<- ground_notes[!duplicated(ground_notes$c_Survey_Notes), ] %>% as_tibble()


kable(filter(ground_notes, Survey_Complete == "No"), caption= paste0("Table 14. Survey Notes recorded during incomplete surveys in ", params$year)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)
  
}
```


#### **Common Eider (*Somateria mollissima*)**

##### Current year (`r params$year`)

```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}

if(nrow(ground_surveys) == 0){
    paste0("No ground-based nest surveys were collected during ", params$year)
    
    } else {

# create table showing current year counts by nest status
kable(ground_methods %>% filter(Species_Code == "COEI"), caption = "Table 15. Nest counts per status category.") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)

# create table showing current year notes by survey status
kable(ground_notes %>% filter(Species_Code == "COEI")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)
}

```

##### Counts over time

```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}

# create violin plot

plt_var(type= "violin", survey = "ground", species = "COEI", var ="Nests")

# plot current year values for each life stage
coei<-filter(nests, Species_Code == "COEI")

PlotBirds(coei, raw_count= TRUE, legend = T, var = "Nests", plot_title = FALSE, overlay = "island", facet = "Species_Code")+
  scale_color_discrete() 
```

#### **Double-crested Cormorant (*Phalacrocorax auritus*)**

##### Current year (`r params$year`)

```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}

if(nrow(ground_surveys) == 0){
    paste0("No ground-based nest surveys were collected during ", params$year)
    
    }else {

# create table showing current year counts by nest status
kable(ground_methods %>% filter(Species_Code == "DCCO"), caption = "Table 18. Nest counts per status category.")%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)

# create table showing current year notes by survey status
kable(ground_notes %>% filter(Species_Code == "DCCO")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)


}


```
\pagebreak 

##### Counts over time

```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}

# create violin plot

plt_var(type= "violin", survey = "ground", species = "DCCO", var ="Nests")


# subset by DCCO to plot
dcco2<-filter(nests, Species_Code == "DCCO")
PlotBirds(dcco2, raw_count= TRUE, legend = TRUE, var = "Nests")
```



#### **Great Black-backed Gull (*Larus marinus*)**

##### Current year (`r params$year`)

```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12, message=FALSE}

if(nrow(ground_surveys) == 0){
    paste0("No ground-based nest surveys were collected during ", params$year)
    
    }else {
      
# create table showing current year counts by nest status
kable(ground_methods %>% filter(Species_Code == "GBBG"), caption = "Table 20. Nest counts per status category.")%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)

# create table showing current year notes by survey status
kable(filter(ground_notes, Species_Code == "GBBG")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)
}

```
\pagebreak 

##### Counts over time

```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}

# create violin plot

plt_var(type= "violin", survey = "ground", species = "GBBG", var ="Nests")


gbbg2<-filter(nests, Species_Code == "GBBG")
PlotBirds(gbbg2, raw_count= TRUE, legend = TRUE, var = "Nests")
```


#### **Herring Gull (*Larus argentatus*)**

##### Current year (`r params$year`)

```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12, message=FALSE}

if(nrow(ground_surveys) == 0){
    paste0("No ground-based nest surveys were collected during ", params$year)
    
    }else {
      
# create table showing current year counts by nest status
kable(ground_methods %>% filter(Species_Code == "HERG"), caption = "Table 22. Nest counts per status category.")%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)

# create table showing current year notes by survey status
kable(filter(ground_notes, Species_Code == "HERG")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)
    }

```
\pagebreak 

##### Counts over time

```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}

# create violin plot

plt_var(type= "violin", survey = "ground", species = "HERG", var ="Nests")

# plot counts over time per island
herg2<-filter(nests, Species_Code == "HERG")
PlotBirds(herg2, raw_count= TRUE, legend = TRUE, var = "Nests")
```



#### **Black-crowned night Heron (*Nycticorax nycticorax*)**

##### Current year (`r params$year`)

```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}

if(nrow(ground_surveys) == 0){
    paste0("No ground-based nest surveys were collected during ", params$year)
    
    }else {
      
# create table showing current year counts by nest status
kable(ground_methods %>% filter(Species_Code == "BCNH"), caption = "Table 24. Nest counts per status category.")%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)

# create table showing current year notes by survey status
kable(filter(ground_notes, Species_Code == "BCNH")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)
}

```
\pagebreak 

##### Counts over time

```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}

# create violin plot

plt_var(type= "violin", survey = "ground", species = "BCNH", var ="Nests")

# plot counts over time per island
bcn0h<-filter(nests, Species_Code == "BCNH")
PlotBirds(bcnh, raw_count= TRUE, legend = TRUE, var = "Nests")
```

#### **Glossy Ibis (*Plegadis falcinellus*)**

##### Current year (`r params$year`)

```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12, message=FALSE}

if(nrow(ground_surveys) == 0){
    paste0("No ground-based nest surveys were collected during ", params$year)
    
    }else {
      
# create table showing current year counts by nest status
kable(ground_methods %>% filter(Species_Code == "GLIB"), caption = "Table 26. Nest counts per status category.")%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)

# create table showing current year notes by survey status
kable(filter(ground_notes,Species_Code == "GLIB")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)
    }

```
\pagebreak 

##### Counts over time

```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}

# create violin plot

plt_var(type= "violin", survey = "ground", species = "GLIB", var ="Nests")

# plot counts over time per island
glib<-filter(nests, Species_Code == "GLIB")
PlotBirds(glib, raw_count= TRUE, legend = TRUE, var = "Nests")
```

#### **Great Egret (*Ardea alba*)**

##### Current year (`r params$year`)

```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}

if(nrow(ground_surveys) == 0){
    paste0("No ground-based nest surveys were collected during ", params$year)
    
    }else {
      
# create table showing current year counts by nest status
kable(ground_methods %>% filter(Species_Code == "GREG"), caption = "Table 28. Nest counts per status category.")%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)

# create table showing current year notes by survey status
kable(filter(ground_notes, Species_Code == "GREG")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)
    }

```
\pagebreak 

##### Counts over time

```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}

# create violin plot

plt_var(type= "violin", survey = "ground", species = "GREG", var ="Nests")

# plot counts over time per island
greg<-filter(nests, Species_Code == "GREG")
PlotBirds(greg, raw_count= TRUE, legend = TRUE, var = "Nests")
```

#### **Snowy Egret (*Egretta thula*)**

##### Current year (`r params$year`)

```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}
if(nrow(ground_surveys) == 0){
    paste0("No ground-based nest surveys were collected during ", params$year)
    
    }else {
      
# create table showing current year counts by nest status
kable(ground_methods %>% filter(Species_Code == "SNEG"), caption = "Table 30. Nest counts per status category.")%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)

# create table showing current year notes by survey status
kable(filter(ground_notes, Species_Code == "SNEG")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)
    }

```

##### Counts over time

```{r,cache=F,echo=F, warning =FALSE, include=T,fig.pos='h',results='asis', fig.height= 8, fig.width= 12,message=FALSE}

# create violin plot

plt_var(type= "violin", survey = "ground", species = "SNEG", var ="Nests")


# plot counts over time per island
sneg<-filter(nests, Species_Code == "SNEG")
PlotBirds(sneg, raw_count= TRUE, legend = TRUE, var = "Nests")
```

<!--

### **Additional Incubation Surveys** {.tabset}



### **Annual Summaries** {.tabset}

#### American Oystercatcher

<--
